### 2016-09-29 paper/abstract.tex
small paper edits

Inference problems are often constrained by state equations, which arise from conservation laws, and constitutive relations. Often, a hierarchy of models can be derived from such laws and relations, reflecting varying fidelity versus cost tradeoffs. We introduce a goal oriented, model adaptive, inference algorithm, which allows users to achieve an aritrary error tolerance, while minmizing the use of high fidelity constraints. Numerical experiments test the algorithm on a highly nonlinear inverse problem, and showcase the computational savings and robustness benefits of this new approach.

>>>>

\red{Inference problems are often constrained by state equations, which arise from conservation laws, and constitutive relations. Often, a hierarchy of models can be derived from such laws and relations, reflecting varying fidelity versus cost tradeoffs. We introduce a goal-oriented, model adaptive, inference algorithm, which allows users to achieve an arbitrary error tolerance, while minimizing the use of high fidelity constraints. Numerical experiments exercise the algorithm on a highly nonlinear inverse problem, and showcase the computational savings and robustness benefits of this new approach.}

### 2016-09-29 paper/algorithms.tex
small paper edits

\section{Goal Oriented Inference Algorithm}\label{sect:alg}
%
Based on the theoretical developments in the last section, we now give a goal-oriented inference algorithm that allows one to combine models of varying fidelity, while maintaining rigorous control of QoI error.

Just as error estimates can be used to guide mesh-refinement~\cite{BecRann01}, the error estimate \cref{eq:finErrExp} can be localized to give elemental contributions and used to guide the division of the domain for a mixed-fidelity model. The error estimate can be calculated again, using the mixed-fidelity model as the lower-fidelity model. This process can be repeated, successively increasing the proportion of the domain in which the high-fidelity model is used, until some threshold is reached. Algorithm~\cref{alg:refSeries} describes this approach. Note that the QoI error adjoint problem \cref{eq:superAdjEq} involves linearization about $\Psi_{HF}$, which is not available, so in the case of a nonlinear goal-oriented inverse problem, the QoI error adjoint problem is approximated by linearizing about $\Psi_{LF}$ instead.
%
\alglanguage{pseudocode}
\begin{algorithm}[h!]
\small
\caption{An algorithm to adaptively build a mixed-fidelity model for low error in the QoI.}
\label{alg:refSeries}
\begin{algorithmic}[1]
\State{Define maximum acceptable absolute relative QoI error \texttt{errTol}}
\State{Define maximum number of adaptive iterations \texttt{maxIter}}
\Procedure{$\texttt{BuildMF}$}{HF model, LF model, \texttt{errTol}, \texttt{maxIter}}
	\State{Let the model MF$_0$ be the LF model applied everywhere in the domain.}
	\State{$i\gets0$}
	\State{Solve for stationary point $\Psi_{MF_0}$ of augmented Lagrangian $\mathcal{M}_{MF_0}$}
	\State{Solve QoI error adjoint equation, linearized about $\Psi_{MF_0}$, for \par\hskip\algorithmicindent supplementary adjoint $\Lambda_0$ (see \cref{eq:superAdjEq})}
	\State{Compute QoI error estimate

@@

%------------------------------------------------------------------------------------------------------------------------%
\subsection{Error Estimate Localization}\label{sec:errLocal}
%------------------------------------------------------------------------------------------------------------------------%
Algorithm~\cref{alg:refSeries} does not require a specific method for localizing the error estimate. A na\"{i}ve approach would be to write the error estimate as a sum of integrals over elements and their boundaries, and calculate the error contribution by each element as the integral over that element. While simple, this method can lead to non-zero error contributions from elements in which the high-fidelity model is already being used, making the error decomposition more difficult to interpret and use for refinement.

>>>>

\section{Goal-Oriented Inference Algorithm}\label{sect:alg}
%
Based on the theoretical developments in the last section, we now give a goal-oriented inference algorithm that allows one to combine models of varying fidelity, while maintaining rigorous control of QoI error.

Just as error estimates can be used to guide mesh-refinement~\cite{BecRann01}, the error estimate \cref{eq:finErrExp} can be localized to give elemental contributions and used to guide the division of the domain for a mixed-fidelity model. The error estimate can be calculated again, using the mixed-fidelity model as the lower-fidelity model. This process can be repeated, successively increasing the proportion of the domain in which the high-fidelity model is used, until some threshold is reached. \Cref{alg:refSeries} describes this approach. Note that the QoI error adjoint problem \cref{eq:superAdjEq} involves linearization about $\Psi_{HF}$, which is not available, so in the case of a nonlinear goal-oriented inverse problem, the QoI error adjoint problem is approximated by linearizing about $\Psi_{LF}$ instead.
%
\alglanguage{pseudocode}
\begin{algorithm}[h!]
\small
\caption{An algorithm to adaptively build a mixed-fidelity model for low error in the QoI.}
\label{alg:refSeries}
\begin{algorithmic}[1]
\State{Define maximum acceptable absolute relative QoI error \texttt{errTol}}
\State{Define maximum number of adaptive iterations \texttt{maxIter}}
\Procedure{$\texttt{BuildMF}$}{HF model, LF model, \texttt{errTol}, \texttt{maxIter}}
	\State{Let the model MF$_0$ be the LF model applied everywhere in the domain.}
	\State{$i\gets0$}
	\State{Solve for stationary point $\Psi_{MF_0}$ of augmented Lagrangian $\mathcal{M}_{MF_0}$}
	\State{Solve QoI error adjoint equation, linearized about $\Psi_{MF_0}$, for \par\hskip\algorithmicindent supplementary adjoint $\Lambda_0$ (see \cref{eq:superAdjEq})}
	\State{Compute QoI error estimate

@@

%------------------------------------------------------------------------------------------------------------------------%
\subsection{Error Estimate Localization}\label{sec:errLocal}
%------------------------------------------------------------------------------------------------------------------------%
\Cref{alg:refSeries} does not require a specific method for localizing the error estimate. A na\"{i}ve approach would be to write the error estimate as a sum of integrals over elements and their boundaries, and calculate the error contribution by each element as the integral over that element. While simple, this method can lead to non-zero error contributions from elements in which the high-fidelity model is already being used, making the error decomposition more difficult to interpret and use for refinement.

### 2016-09-29 paper/intro.tex
small paper edits

In this work, we extend the use of multi-fidelity models to inverse problems. Specifically, we consider the case where there is a target quantity of interest to be computed from the inferred parameters. The use of concurrent multi-fidelity models for the goal-oriented forward problem is described in \cite{OdenPrudetal06}, a method using adjoints for the goal-oriented forward problem is described; an extension to inverse problems is described in \cite{OdenPrudetal10}.\red{\footnote{Vikram, this was known at the time of the thesis, but not not cited in the thesis. I'm not actually sure how to handle this, since this tries to solve the same problem we do, but I don't understand it well enough to do a detailed analysis. I did manage to find the book, if you'd like me to scan the pages to you...}} We present a systematic method to use different constraints in different regions of the domain, while controlling the error due to the use of lower fidelity models, in a target quantity of interest. Our approach is based on a novel use of adjoint error estimation, which enables the computation of third-order QoI error estimates, at the cost of solving an additional adjoint problem.

The rest of the paper is organized as follows, \cref{sect:form} presents the mathematical formulation and error analysis for the multi-model inverse problem. Then, in \cref{sect:alg} we discuss the goal oriented inference algorithm and present a computational complexity analysis. \cref{sect:numexp} shows the application of this algorithm to a model problem and a contaminant flow problem. Finally, we present conclusions and direction for future work in \cref{sect:conc}.

%% \subsection{draft outline}
%% \bit
%% \item background/previous work? to put this work in context?
%%   %this is probably enough? at least for previous work...
%%   %what context though? is this another way to approach inverse problems?
%% \item mathematical formulation
%%   \bit
%%   \item define inverse problem and qoi
%%   \item derive error estimate
%%   \item caveats (models to which this can be applied, no guarantee that mixed model will be worth the cost to obtain, no bounds on accuracy of estimate in nonlinear case)
%%   \item relation to mesh refinement
%%   \item How does this compare to just using Oden et al's method (multimodel, foward with QoI) and replacing forward PDEs with optimality system of inverse problem? (Oden's gets you second-order in error psiHF-psiLF, but this one gets third, though you have to compute the super-adjoint...linear solve, though...big system, but linear...)
%%     %solving for adjoint and forward simultaneously (KKT) might cause you to lose an order of accuracy?
%%     %where was that book that Oden might have mentioned inverse application of his method?
%%       %Multiscale methods : bridging the scales in science and engineering / edited by Jacob Fish. (TA335.M85 2010 in barker)
%%   \eit
%% \item numerical results (perhaps not include all of these...)
%%   \bit
%%   \item convection-diffusion(-reaction)
%%     \bit
%%     \item interpretation of error estimate breakdown (interaction of QoI with observations)
%%     \item large reaction term: MF model more cooperative with nonlinear solver
%%     \eit
%%   \item convection-diffusion-reaction, field vs scalar description of parameters
%%     \bit
%%     \item interface question: do we enforce continuity between the two models' parameters?
%%       %-navierstokes+stokes seems to require continuity of flux and velocity across boundary...?
%%       %-is it okay to mix diff and convdiff for the forward like we did? does the discontinuous velocity break any intergration-by-parts   assumptions? (no that's only if discontinuity in diffusion bit...)
%%     \item chose to have one 'scalar' parameter, so all sections where the LF used shared same single parameter; is this necessarily the best choice?
%%     %alternative way to describe 'scalar' case? but here we say there is one scalar, so this would be how to combine them...if you wanted plateaus to be different, can let LF model be two-scalars...
%%     %fig 3.10 -> what does the inferred field look like? does it look like two different plateaus?
%%       % practice/T_channel/diff_param_res/with_reaction/long_channel_stash/qoi3_setup02_r4p2_deref/ (pushed 10/3/15)
%%         %at the very last refinement, most of refined region seems to be two large humps...and a section that looks pretty much like the surrounding scalar...so maybe 3 patches...
%%     \item chose to enforce continuity (easier to implement; field representation required to be continuous); algorithm wanted 60/40 HF/LF mix
%%     \eit
%%   \item alternate/additional example: use setup from B+V? navier-stokes and another model?
%%     \bit
%%     \item potential flow? model flow as sum of elementary flows; but then parameters would be magnitudes of elementary flows, which has no spatial location...is that a problem for the algorithm or just for interpreting error breakdown?
%%     \item there would still be the question of how to interface the models...(in stokes vs navier-stokes foward paper, they just required continuity of state variables (velocity field and presssure))
%%     %do scalar + field for infer BCs on BV NS thing? try for different QoIs? (can we have a variable live on a boundary while the rest are inside? vikram will poke roy about implementing)
%%     %stokes is good in slow flows...not in boundary layer cuz gradient is high...
%%     %what situation would it make sense to use the LF model in some place?
%%     \eit
%%   \item have a separate section to address questions of how to mix models? (interfaces, whether 'scalar' parameter should be the same everywhere)
%%   \eit
%% \item conclusion...what goes in this section usually? summary and directions for future work?
%% \eit

>>>>

In this work, we extend the use of multi-fidelity models to inverse problems. Specifically, we consider the case where there is a target quantity of interest to be computed from the inferred parameters. The use of concurrent multi-fidelity models for the goal-oriented forward problem is described in \cite{OdenPrudetal06}, a method using adjoints for the goal-oriented forward problem is described; an extension to inverse problems is described in \cite{OdenPrudetal10}. We present a systematic method to use different constraints in different regions of the domain, while controlling the error due to the use of lower fidelity models, in a target quantity of interest. Our approach is based on a novel use of adjoint error estimation, which enables the computation of third-order QoI error estimates, at the cost of solving an additional adjoint problem.

The rest of the paper is organized as follows. \Cref{sect:form} presents the mathematical formulation and error analysis for the multi-model inverse problem. Then, in \Cref{sect:alg} we discuss the goal-oriented inference algorithm and present a computational complexity analysis. \Cref{sect:numexp} shows the application of this algorithm to a model problem and a contaminant flow problem. Finally, we present conclusions and directions for future work in \Cref{sect:conc}.

### 2016-09-29 paper/numerical_experiments.tex
small paper edits

\section{Numerical Experiments}\label{sect:numexp}
%
We now use \cref{alg:refSeries} to solve goal oriented inverse problems in a multi-model setting; the method used to localize the error estimate is described in \cref{sec:errLocal}. We first consider simple two-dimensional models in order to explore the behavior of the algorithm in relation to iteration number as well as the placement of observations and QoI regions. In \cref{sec:cdvcdr}, the high fidelity model is a convection-diffusion-reaction nonlinear model, and the low fidelity model is a linear convection-diffusion model. We apply \cref{alg:refSeries} to this pair of models, and then examine how the localized error estimate is affected by changes in sensor placement and in the QoI region. Finally, in \cref{sec:diffvcdr3D}, we consider a more realistic pair of three-dimensional models to illustrate computational savings \red{when using \cref{alg:refSeries}, as compared to solving the high-fidelity inverse problem.}\footnote{\red{Well, compuational savings are only seen in the online-offline case, so that's not the referenced algorithm per se...}}

@@

%------------------------------------------------------------------------------------------------------------------------%
\subsection{Convection-Diffusion(-Reaction)} \label{sec:cdvcdr}
%------------------------------------------------------------------------------------------------------------------------%
In this section, we consider a pair of models which differ in the physics included. In \cref{sec:cdvcdrSetup} we describe a baseline setup for a simple two-dimensional problem. \cref{sec:cdvcdrBaseRef} describes the results of applying \cref{alg:refSeries} to the baseline problem, and \cref{sec:qoivdata} describes the results of changing the placement of the observations or the QoI region from the baseline.
%
%------------------------------------------------------------%
\subsubsection{Problem Setup} \label{sec:cdvcdrSetup}
%------------------------------------------------------------%
%
We consider a rectangular domain $\Omega(x_1,x_2)=[0,5]\times[0,1]$, where $x_1$ and $x_2$ are the spatial coordinates. The high-fidelity model is a single-species convection-diffusion-reaction equation with a nonlinear reaction term, described by,
%
\begin{subequations}
\label{eq:cdvcdrHF}
\begin{align}
k_d\nabla^2 u - \vec{V}\cdot\nabla u + k_ru^2 = f(q) \quad &\text{in } \Omega, \label{eq:cdvcdrHF_int} \\
u = 0 \quad &\text{on } \partial \Omega \label{eq:cdvcdrHF_bdry}
\end{align} 
\end{subequations}
%
where the state $u$ is the species concentration and $f(q)$ is a forcing field described by the parameters. We have a divergence-free parabolic-profile velocity field $\vec{V}(x_1,x_2) = (2x_2(1-x_2),0)$; the diffusion and reaction coefficients are $k_d = 0.1$ and $k_r = -42.0$, respectively. The low-fidelity model,
%
\begin{equation}
k_d\nabla^2 u - \vec{V}\cdot\nabla u = f(q)
\end{equation}
%
differs only in the removal of the reaction term. To form the mixed-fidelity models, we divide the domain into complementary subdomains, $\Omega_{HF}$ and $\Omega_{LF}$, where the high- and low-fidelity models are solved, respectively. The resulting mixed-fidelity models can be described by, 
%
\begin{equation}
k_d\nabla^2 u - \vec{V}\cdot\nabla u + k^{MF}_ru^2= f(q),
\end{equation}
%
where $k^{MF}_r$ is a piecewise-constant reaction coefficient,
%
\begin{equation}
k^{MF}_r=
\begin{cases}
-42.0 & \textrm{if }x\in\Omega_{HF} \\
0 & \textrm{if }x\in\Omega_{LF}.
\end{cases}
\end{equation}
%
Homogeneous Dirichlet boundary conditions are applied on the entire boundary of the domain. The QoI we wish to calculate is the integral of the state,
%
\begin{equation}
I(q,u)=\int_{(x_1,x_2)\in \Omega_I} u \:\textrm{d}A,
\end{equation}
%
over a region $\Omega_I=[0.625,0.875]\times[0.375,0.625]$.

The unknown parameters we wish to infer correspond to the forcing field, so that $f(q)=q$. For the low-fidelity model, the inverse problem is linear (the inferred parameters are linear in the observations). Observations consisting of the state at three points in the domain are artificially generated by running the high-fidelity model on a finer mesh with the true forcing field
%
\begin{equation}
f_{true}(x_1,x_2)=
\begin{cases}
1.0 & \textrm{if }(x_1,x_2)\in[0.125,0.375]\times[0.125,0.375] \\
0.8 & \textrm{if }(x_1,x_2)\in[2.375,2.625]\times[0.375,0.625] \\
0 & \textrm{otherwise}.
\end{cases}
\end{equation}
%
The locations of the observations and the region $\Omega_I$ over which the QoI is calculated are shown in \cref{fig:baseSetup}. Since the inverse problem is ill-posed, we use Tikhonov regularization~\cite{EngHanNeu00}; the regularization term is $\frac{\beta}{2}\int_\Omega \|\nabla f(q)\|_2^2\:\textrm{d}A$, where $\beta=10^{-5}$ is the regularization coefficient. 
%
\begin{figure}[htbp]
\centering
\includegraphics[width=0.8\textwidth]{baseSeries/setup_3_3.pdf}
\caption{Locations of the observations and the QoI region.}
\label{fig:baseSetup}
\end{figure}
%

For the numerical simulations, we use the finite element method (FEM), employing a continuous Galerkin formulation with Lagrange elements. We use the \texttt{libMesh} library~\cite{libMeshPaper} for the FEM calculations. 
%\red{The library offers easy calculation of adjoint systems, error estimates and subdomain restricted variables.}\footnote{But we don't use Libmesh's error estimates or restrict variables to subdomains...} 
The domain is discretized by a regular mesh of quadrilaterals, with 250 and 50 elements along the $x_1$ and $x_2$ directions, respectively, for a total of 12,500 elements, resulting in 12,801 degrees of freedom per variable. The diffusion coefficient is chosen so that the cell P\'{e}clet number never exceeds 0.1, and thus no stabilization is required.
%
%------------------------------------------------------------%
\subsubsection{Adaptive Model Refinement Results} \label{sec:cdvcdrBaseRef} 
%------------------------------------------------------------%
%

We now present the results for solving the inference problem using \cref{alg:refSeries}. Once the QoI error estimate is calculated using \ref{eq:finErrExp}, the error estimate is then decomposed into local contributions, as described in \cref{eq:basisblame}. At each iteration, based on this decomposition, we choose the basis functions with the largest error contributions until an additional 5\% of the elements has been marked for refinement. This is repeated until the estimated absolute relative error in the QoI, calculated as $\epsilon_i/(\epsilon_i+I(q_{MF_i},u_{MF_i}))$, is less than $1\%$.

@@

\Cref{fig:baseErr} shows the true and estimated absolute relative errors in the QoI for the various mixed-fidelity models generated by \cref{alg:refSeries}; the true and estimated relative errors are calculated relative to the true and estimated high-fidelity QoI, respectively. In this case, we see that the QoI error of only $1\%$ is attained with a mixed-fidelity model where the high-fidelity model is used in only about $10\%$ of the domain. We note that, in general, there is no guarantee that either the error in the QoI or the relative error in the error estimate will decrease monotonically as more of the domain is refined.
%
\begin{figure}[htbp]
\centering
\includegraphics[width=0.8\textwidth]{baseSeries/err_est.pdf}
\caption{True and estimated absolute relative error in QoI, plotted as a function of the percentage area of the domain in which the high-fidelity convection-diffusion-reaction model is used. \red{Does this even have enough points to warrant a plot? table instead?}}
\label{fig:baseErr}
\end{figure}
%

@@

%
\begin{table}
\caption{Average times to solve inverse problem and obtain error estimate with mixed- and high-fidelity models and noisy datasets.}
\label{tab:ref3D_newdata_times_diffmesh}
\centering
\begin{tabular}{ccc|c|c|c}
\cline{4-5} 
 & & & \multicolumn{2}{|c|}{Error Estimation} & \\
\cline{1-6}
\multicolumn{1}{|c|}{\multirow{3}{*}{Case}} & \multicolumn{1}{|c|}{\multirow{3}{*}{DOFs}} & Inverse & Auxiliary & Supplementary & \multicolumn{1}{|c|}{Total} \\
\multicolumn{1}{|c|}{} & \multicolumn{1}{|c|}{} & Problem & Variables & Adjoint & \multicolumn{1}{|c|}{Solution}\\
\multicolumn{1}{|c|}{} & \multicolumn{1}{|c|}{} & Time (s) &  Time (s) & Time (s) & \multicolumn{1}{|c|}{Time (s)}\\
\cline{1-6}
\multicolumn{1}{|c|}{LF}    & \multicolumn{1}{|c|}{9537}   & 16   & --  & -- & \multicolumn{1}{|c|}{--} \\ \hline
\multicolumn{1}{|c|}{MF01}  & \multicolumn{1}{|c|}{13147}  & 185  & 107 & 235 & \multicolumn{1}{|c|}{526} \\ \hline
\multicolumn{1}{|c|}{MF02}  & \multicolumn{1}{|c|}{17895}  & 328  & 169 & 206 & \multicolumn{1}{|c|}{703} \\ \hline
\multicolumn{1}{|c|}{MF03}  & \multicolumn{1}{|c|}{21001}  & 435  & 202 & 185 & \multicolumn{1}{|c|}{821} \\ \hline
\multicolumn{1}{|c|}{MF04}  & \multicolumn{1}{|c|}{24528}  & 406  & 201 & 188 & \multicolumn{1}{|c|}{795} \\ \hline
\multicolumn{1}{|c|}{MF05}  & \multicolumn{1}{|c|}{27984}  & 498  & 263 & 198 & \multicolumn{1}{|c|}{959} \\ \hline
\multicolumn{1}{|c|}{HF}    & \multicolumn{1}{|c|}{70785}  & 1185 & --  & --  & \multicolumn{1}{|c|}{1185} \\ \hline
\end{tabular}
\end{table}
%
\begin{table}
\caption{Average times to solve inverse problem and obtain error estimate with mixed- and high-fidelity models and noisy datasets; all averages only over those datasets for which the high-fidelity problem converged when given the low-fidelity problem as an intial guess.}
%\label{tab:ref3D_newdata_times_diffmesh}
\centering
\begin{tabular}{ccc|c|c|c}
\cline{4-5} 
 & & & \multicolumn{2}{|c|}{Error Estimation} & \\
\cline{1-6}
\multicolumn{1}{|c|}{\multirow{3}{*}{Case}} & \multicolumn{1}{|c|}{\multirow{3}{*}{DOFs}} & Inverse & Auxiliary & Supplementary & \multicolumn{1}{|c|}{Total} \\
\multicolumn{1}{|c|}{} & \multicolumn{1}{|c|}{} & Problem & Variables & Adjoint & \multicolumn{1}{|c|}{Solution}\\
\multicolumn{1}{|c|}{} & \multicolumn{1}{|c|}{} & Time (s) &  Time (s) & Time (s) & \multicolumn{1}{|c|}{Time (s)}\\
\cline{1-6}
\multicolumn{1}{|c|}{LF}    & \multicolumn{1}{|c|}{9537}   & 16   & --  & -- & \multicolumn{1}{|c|}{--} \\ \hline
\multicolumn{1}{|c|}{MF01}  & \multicolumn{1}{|c|}{13147}  & 156  & 101 & 261 & \multicolumn{1}{|c|}{517} \\ \hline
\multicolumn{1}{|c|}{MF02}  & \multicolumn{1}{|c|}{17895}  & 304  & 175 & 183 & \multicolumn{1}{|c|}{662} \\ \hline
\multicolumn{1}{|c|}{MF03}  & \multicolumn{1}{|c|}{21001}  & 385  & 228 & 187 & \multicolumn{1}{|c|}{801} \\ \hline
\multicolumn{1}{|c|}{MF04}  & \multicolumn{1}{|c|}{24528}  & 347  & 202 & 193 & \multicolumn{1}{|c|}{743} \\ \hline
\multicolumn{1}{|c|}{MF05}  & \multicolumn{1}{|c|}{27984}  & 470  & 275 & 212 & \multicolumn{1}{|c|}{957} \\ \hline
\multicolumn{1}{|c|}{HF}    & \multicolumn{1}{|c|}{70785}  & 1185 & --  & --  & \multicolumn{1}{|c|}{1185} \\ \hline
\end{tabular}
\end{table}
%

>>>>

\section{Numerical Experiments}\label{sect:numexp}
%
We now use \cref{alg:refSeries} to solve goal-oriented inverse problems in a multi-model setting; the method used to localize the error estimate is described in \cref{sec:errLocal}. We first consider simple two-dimensional models in order to explore the behavior of the algorithm as well as the effect of varying the placement of observations and QoI regions. In \cref{sec:cdvcdr}, the high fidelity model is a convection-diffusion-reaction nonlinear model, and the low fidelity model is a linear convection-diffusion model. We apply \cref{alg:refSeries} to this pair of models, and then examine how the localized error estimate is affected by changes in sensor placement and in the QoI region. In \cref{sec:diffvcdr3D}, we consider a more realistic pair of three-dimensional models.

@@

%------------------------------------------------------------------------------------------------------------------------%
\subsection{Convection-Diffusion(-Reaction)} \label{sec:cdvcdr}
%------------------------------------------------------------------------------------------------------------------------%
In this section, we consider a pair of models which differ in the physics included. In \cref{sec:cdvcdrSetup} we describe a baseline setup for a simple two-dimensional problem. \Cref{sec:cdvcdrBaseRef} describes the results of applying \cref{alg:refSeries} to the baseline problem, and \cref{sec:qoivdata} describes the results of changing the placement of the observations or the QoI region from the baseline.
%
%------------------------------------------------------------%
\subsubsection{Problem Setup} \label{sec:cdvcdrSetup}
%------------------------------------------------------------%
%
We consider a rectangular domain $\Omega(x_1,x_2)=[0,5]\times[0,1]$, where $x_1$ and $x_2$ are the spatial coordinates. The high-fidelity model is a single-species convection-diffusion-reaction equation with a nonlinear reaction term, described by,
%
\begin{subequations}
\label{eq:cdvcdrHF}
\begin{align}
k_d\nabla^2 u - \vec{V}\cdot\nabla u + k_ru^2 = f(q) \quad &\text{in } \Omega, \label{eq:cdvcdrHF_int} \\
u = 0 \quad &\text{on } \partial \Omega \label{eq:cdvcdrHF_bdry}
\end{align} 
\end{subequations}
%
where the state $u$ is the species concentration and $f(q)$ is a forcing field described by the parameters. We have a divergence-free parabolic-profile velocity field $\vec{V}(x_1,x_2) = (2x_2(1-x_2),0)$; the diffusion and reaction coefficients are $k_d = 0.1$ and $k_r = -42.0$, respectively. The low-fidelity model,
%
\begin{equation}
k_d\nabla^2 u - \vec{V}\cdot\nabla u = f(q)
\end{equation}
%
differs only in the removal of the reaction term. To form the mixed-fidelity models, we divide the domain into complementary subdomains, $\Omega_{HF}$ and $\Omega_{LF}$, where the high- and low-fidelity models are solved, respectively. The resulting mixed-fidelity models can be described by, 
%
\begin{equation}
k_d\nabla^2 u - \vec{V}\cdot\nabla u + k^{MF}_ru^2= f(q),
\end{equation}
%
where $k^{MF}_r$ is a piecewise-constant reaction coefficient,
%
\begin{equation}
k^{MF}_r=
\begin{cases}
k_r & \textrm{if }x\in\Omega_{HF} \\
0 & \textrm{if }x\in\Omega_{LF}.
\end{cases}
\end{equation}
%
Homogeneous Dirichlet boundary conditions are applied on the entire boundary of the domain. The QoI we wish to calculate is the integral of the state,
%
\begin{equation}
I(q,u)=\int_{(x_1,x_2)\in \Omega_I} u \:\textrm{d}A,
\end{equation}
%
over a region $\Omega_I=[0.625,0.875]\times[0.375,0.625]$.

The unknown parameters we wish to infer correspond to the forcing field, so that $f(q)=q$. For the low-fidelity model, the inverse problem is linear (the inferred parameters are linear in the observations). Synthetic observations consisting of the state at three points in the domain are artificially generated by running the high-fidelity model on a finer mesh with the true forcing field
%
\begin{equation}
f_{true}(x_1,x_2)=
\begin{cases}
1.0 & \textrm{if }(x_1,x_2)\in[0.125,0.375]\times[0.125,0.375] \\
0.8 & \textrm{if }(x_1,x_2)\in[2.375,2.625]\times[0.375,0.625] \\
0 & \textrm{otherwise}.
\end{cases}
\end{equation}
%
The locations of the observations and the region $\Omega_I$ over which the QoI is calculated are shown in \cref{fig:baseSetup}. Since the inverse problem is ill-posed, we use Tikhonov regularization~\cite{EngHanNeu00}; the regularization term is $\frac{\beta}{2}\int_\Omega \|\nabla f(q)\|_2^2\:\textrm{d}A$, where $\beta=10^{-5}$ is the regularization coefficient. 
%
\begin{figure}[htbp]
\centering
\includegraphics[width=0.8\textwidth]{baseSeries/setup_3_3.pdf}
\caption{Locations of the observations and the QoI region.}
\label{fig:baseSetup}
\end{figure}
%

For the numerical simulations, we use the finite element method (FEM), employing a continuous Galerkin formulation with Lagrange elements. We use the \texttt{libMesh} library~\cite{libMeshPaper} for the FEM calculations. 
%\red{The library offers easy calculation of adjoint systems, error estimates and subdomain restricted variables.}\footnote{But we don't use Libmesh's error estimates or restrict variables to subdomains...} 
The domain is discretized by a regular mesh of quadrilaterals, with 250 and 50 elements along the $x_1$ and $x_2$ directions, respectively, for a total of 12,500 elements, resulting in 12,801 degrees of freedom per variable. The diffusion coefficient is such that the cell P\'{e}clet number never exceeds 0.1, and thus no stabilization is required.
%
%------------------------------------------------------------%
\subsubsection{Adaptive Model Refinement Results} \label{sec:cdvcdrBaseRef} 
%------------------------------------------------------------%
%

We now present the results for solving the inference problem using \cref{alg:refSeries}. Once the QoI error estimate is calculated using \cref{eq:finErrExp}, the error estimate is then decomposed into local contributions, as described in \cref{eq:basisblame}. At each iteration, based on this decomposition, we choose the basis functions with the largest error contributions until an additional 5\% of the elements has been marked for refinement. This is repeated until the estimated absolute relative error in the QoI, calculated as $\epsilon_i/(\epsilon_i+I(q_{MF_i},u_{MF_i}))$, is less than $1\%$.

@@

\Cref{fig:baseErr} shows the true and estimated absolute relative errors in the QoI for the various mixed-fidelity models generated by \cref{alg:refSeries}; the true and estimated relative errors are calculated relative to the true and estimated high-fidelity QoI, respectively. In this case, we see that the QoI error of only $1\%$ is attained with a mixed-fidelity model where the high-fidelity model is used in only about $10\%$ of the domain. We note that, in general, there is no guarantee that either the error in the QoI or the relative error in the error estimate will decrease monotonically as more of the domain is refined.
%
\begin{figure}[htbp]
\centering
\includegraphics[width=0.8\textwidth]{baseSeries/err_est.pdf}
\caption{True and estimated absolute relative error in QoI, plotted as a function of the percentage area of the domain in which the high-fidelity convection-diffusion-reaction model is used.}
\label{fig:baseErr}
\end{figure}
%

@@

%
\begin{table}
\caption{Average times to solve inverse problem and obtain error estimate with mixed- and high-fidelity models and noisy datasets.}
\label{tab:ref3D_newdata_times_diffmesh}
\centering
\begin{tabular}{ccc|c|c|c}
\cline{4-5} 
 & & & \multicolumn{2}{|c|}{Error Estimation} & \\
\cline{1-6}
\multicolumn{1}{|c|}{\multirow{3}{*}{Case}} & \multicolumn{1}{|c|}{\multirow{3}{*}{DOFs}} & Inverse & Auxiliary & Supplementary & \multicolumn{1}{|c|}{Total} \\
\multicolumn{1}{|c|}{} & \multicolumn{1}{|c|}{} & Problem & Variables & Adjoint & \multicolumn{1}{|c|}{Solution}\\
\multicolumn{1}{|c|}{} & \multicolumn{1}{|c|}{} & Time (s) &  Time (s) & Time (s) & \multicolumn{1}{|c|}{Time (s)}\\
\cline{1-6}
\multicolumn{1}{|c|}{LF}    & \multicolumn{1}{|c|}{9537}   & 16   & --  & -- & \multicolumn{1}{|c|}{--} \\ \hline
\multicolumn{1}{|c|}{MF01}  & \multicolumn{1}{|c|}{13147}  & 185  & 107 & 235 & \multicolumn{1}{|c|}{526} \\ \hline
\multicolumn{1}{|c|}{MF02}  & \multicolumn{1}{|c|}{17895}  & 328  & 169 & 206 & \multicolumn{1}{|c|}{703} \\ \hline
\multicolumn{1}{|c|}{MF03}  & \multicolumn{1}{|c|}{21001}  & 435  & 202 & 185 & \multicolumn{1}{|c|}{821} \\ \hline
\multicolumn{1}{|c|}{MF04}  & \multicolumn{1}{|c|}{24528}  & 406  & 201 & 188 & \multicolumn{1}{|c|}{795} \\ \hline
\multicolumn{1}{|c|}{MF05}  & \multicolumn{1}{|c|}{27984}  & 498  & 263 & 198 & \multicolumn{1}{|c|}{959} \\ \hline
\multicolumn{1}{|c|}{HF}    & \multicolumn{1}{|c|}{70785}  & 1185 & --  & --  & \multicolumn{1}{|c|}{1185} \\ \hline
\end{tabular}
\end{table}
%
We see that the mixed-fidelity models, when applied to noisy datasets different from that with which they were generated, continue to perform well in achieving a small error in the QoI while limiting the use of the high-fidelity model to less than a third of the domain. Given the same initial guess, the mixed-fidelity inverse problems takes less time on average than the high-fidelity inverse problem to solve, and they converge consistently.

In testing this online-offline setting with a larger range of datasets, we would like to note some additional observations. Suppose one has two datasets $d_1$ and $d_2$, the latter of which one wishes to perform inference. For a fixed proportion of domain refinement, a mixed-fidelity model that was adaptively generated $d_2$ gives a smaller QoI error than a mixed-fidelity model that was generated based on $d_1$. The performance gap tends to widen as the two datasets become more different. Suppose one generated $d_1$ using the high-fidelity model $d_2$ using a model which differed from the high-fidelity model in its domain, boundary conditions, and reaction coefficient; then a mixed-fidelity model adaptively formed based on $d_1$ might not give a low QoI error when used to infer from $d_2$. However, when the adaptivity algorithm is directly applied to this $d_2$, it is still able to find a mixed-fidelity model with a low QoI error. This behavior suggest that, if resources allow, it is best to perform the adaptivity online with the true observations. If one is confident that one's high-fidelity model is a good approximation of reality, then the online-offline approach can potentially give a faster yet accurate QoI estimate.
